Software Citation Principles

### 2016-03-18 sc-principles/software-citation-principles.tex
updates in unique identification discussion is section 5

\begin{enumerate}
\item \textbf{Importance}: \label{principle:importance}
Software should be considered a legitimate, citable product of research. Software citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications and data. Software should be cited whenever and wherever a research product (such as a paper or derived software) relies upon it, specifically, as part of the standard reference list for that research product. 
\item \textbf{Credit and Attribution}:
Software citations should facilitate giving scholarly credit and normative
and legal attribution to all contributors to the software, recognizing
that a single style or mechanism of attribution may not be applicable to
all software. 
\item \textbf{Unique Identification}:
A software citation should include a method for identification that is
machine actionable, globally unique, interoperable, and recognized by
a community.
\item \textbf{Persistence}: \label{principle:persistence}
Unique identifiers and metadata describing the software and its disposition should persist~-- even beyond the lifespan of the software they describe.
\item \textbf{Accessibility}: \label{principle:accessibility}
Software citations should permit and facilitate access to the software itself and to such associated metadata, documentation, data, and other materials as are necessary for both humans and machines to make informed use of the referenced software. 
\item \textbf{Specificity}:
Software citations should facilitate identification of, and access to, the specific version of software that was used.  Software identification should be as specific as necessary, such as using version numbers, revision numbers, or variants such as platforms.
%\item \textbf{Interoperability and Flexibility}:
%Software citation methods should be sufficiently flexible to accommodate the variant practices among communities, but should not differ so much that they compromise interoperability of software citation practices across communities.
\end{enumerate}

@@

\textbf{Unique identifiers}

The ``unique'' in a UID means that it points to a unique, specific software. However, multiple UIDs might point to the same software.  This is not recommended, but is possible.  We strongly recommend that if there is already a UID for a version of software, no additional UID should be
created.  Multiple UIDs can lead to split credit, which does not help anyone.  In particular, we
urge the use of standard DOIs as the UIDs for software. \katznote{This topic is discussed in detail in the `Unique Identification' section of \cite{10.7717/peerj-cs.1}}

>>>>

\begin{enumerate}
\item \textbf{Importance}: \label{principle:importance}
Software should be considered a legitimate, citable product of research. Software citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications and data. Software should be cited whenever and wherever a research product (such as a paper or derived software) relies upon it, specifically, as part of the standard reference list for that research product. 
\item \textbf{Credit and Attribution}: \label{principle:credit}
Software citations should facilitate giving scholarly credit and normative
and legal attribution to all contributors to the software, recognizing
that a single style or mechanism of attribution may not be applicable to
all software. 
\item \textbf{Unique Identification}: \label{principle:uid}
A software citation should include a method for identification that is
machine actionable, globally unique, interoperable, and recognized by
a community.
\item \textbf{Persistence}: \label{principle:persistence}
Unique identifiers and metadata describing the software and its disposition should persist~-- even beyond the lifespan of the software they describe.
\item \textbf{Accessibility}: \label{principle:accessibility}
Software citations should permit and facilitate access to the software itself and to such associated metadata, documentation, data, and other materials as are necessary for both humans and machines to make informed use of the referenced software. 
\item \textbf{Specificity}:
Software citations should facilitate identification of, and access to, the specific version of software that was used.  Software identification should be as specific as necessary, such as using version numbers, revision numbers, or variants such as platforms.
%\item \textbf{Interoperability and Flexibility}:
%Software citation methods should be sufficiently flexible to accommodate the variant practices among communities, but should not differ so much that they compromise interoperability of software citation practices across communities.
\end{enumerate}

@@

\textbf{Unique identification}

The Unique identification principle (\ref{principle:uid}) calls for ``a method for identification that is machine actionable, globally unique, interoperable, and recognized by a community.''
What this means for data is discussed in detail in the `Unique Identification' section of a report by the FORCE11 Data Citation Implementation Group (DCIG)~\cite{10.7717/peerj-cs.1}, which calls for ``unique identification in a manner that is machine-resolvable on the Web and demonstrates a long-term commitment to persistence.''
This report also lists examples of identifiers that match these criteria including DOIs, PURLs, Handles, ARKS, and NBNs.
For software, we recommend the use of DOIs as the unique identifier due to their common usage and acceptance, particularly as they are the standard for other digital products such as publications.

Note that the ``unique'' in a UID means that it points to a unique, specific software. However, multiple UIDs might point to the same software.
This is not recommended, but is possible.
We strongly recommend that if there is already a UID for a version of software, no additional UID should be created.
Multiple UIDs can lead to split credit, which goes against the Credit and Attribution principle (\ref{principle:credit}).
